Mapping Cancer: the Potential of Cartograms and Alternative Map Displays

### 2020-05-26 letter2/submission2.Rmd

**Reviewer suggestion:**
**Can you clarify whether this recommendation holds when using a ratio measure – which has 1 as the average, and then diverging values higher and lower than the average (eg an SIR).**
Diverging colour schemes have been suggested for this use by Cynthia Brewer, however a ratio measure requires rescaling of the colour space:

When communicating information using colour, a map creator should use a scheme with perceptually uniform color spaces that match equal steps in data space with equal steps in the colour space [@PUCS], a linear color gradient is appropriate for incidence counts and rates, but may not be appropriate for ratio measures.

@@

**Reviewer suggestion:** XXX
I would suggest replacing this section with a discussion of relative versus absolute measures.

@@

We thank you for your consideration of our changes.

>>>>

**Reviewer suggestion:**
**Can you clarify whether this recommendation holds when using a ratio measure – which has 1 as the average, and then diverging values higher and lower than the average (eg an SIR).**
A linear colour gradient can be transformed to a log scale for ratio measures. This has been used in the Australian Cancer Atlas, a small example has been included in the paragraph to explain the transformation.

@@

**Reviewer suggestion:**
**I would suggest replacing this section with a discussion of relative versus absolute measures.**
We have included additional comments discussing of relative measures.

Please pass our warm thanks to the reviewers for their thorough and thoughtful recommendations. Their input has greatly improved this piece of writing.

@@

We thank you for your patience and your consideration of our changes.

### 2020-05-26 literature.Rmd

# 1. Introduction
<!-- 436 words -->
<!-- What data are we working with, why, what are the purposes -->
Researchers, health authorities, governments, not-for-profits and the media are common communicators of cancer statistics. They often present statistics to the public as aggregated values for geopolitical areas. Presenting these statistics requires aggregating individual observations for the geographical units, especially for privacy protection, but also for political and policy purposes. Examples of typical geographical units include states, provinces, local government areas, and post/zip codes. Counts and incidence rates are often obtained by state health registries. For this data, privacy and ethics are key concerns, even when combined to small area statistics.
This type of data is routinely collected for public health reasons and may be made available to the general public as a service to the community.

@@

<!-- statistics -->
Many statistics are commonly used in cancer displays. It is common to see incidence rates, or relative rates which measure how far a region is above or below the average. The purpose of using an age standardized relative rate is, perhaps the desire to pinpoint the areas that need attention because they have higher than expected rates. 
<!--However, we lose the incidence rate information and thus interpretability. --> A region might be much higher than average, but it may not be close to a health concern, because all regions have a low incidence. Supplementary materials can allow map users to recognise when this occurs.

>>>>

# 1. Introduction
<!-- 436 words -->
<!-- What data are we working with, why, what are the purposes -->

Researchers, health authorities, governments, not-for-profits and the media are common communicators of cancer statistics. They often present statistics to the public as aggregated values for geopolitical areas. Counts and incidence rates are often obtained by state health registries. Data privacy and ethics are key concerns even when combined to small area statistics. Presenting these statistics requires aggregating individual observations for the geographical units, also for political and policy purposes. Examples of typical geographical units include states, provinces, local government areas, and post/zip codes. 
This type of data is routinely collected for public health reasons and may be made available to the general public as a service to the community.

@@

Many statistics are commonly used in cancer displays. It is common to see incidence rates, or ratios which displays how far a region is above or below the average. The purpose of using an age standardized ratio is, perhaps the desire to pinpoint the areas that need attention because they have higher than expected rates. Ratios may negatively impact the interpretability of the actual rates of incidence or mortality. However, this impact is offset by the value of seeing the relativity of the values in reference to the mean, especially when the mean value is also given on the display. This helps to put individual areas in perspective, 
as a region might have a value higher than the average, but it may not be a health concern if all regions have a low incidence rate. Supplementary materials such as displaying the mean, or a distribution plot, can allow map users to recognise when this occurs.
